---
title: "Fast Statistical Hypothesis Tests using fastats"
output: 
  html_document:
    code_folding: show
    theme: cosmo
vignette: >
  %\VignetteIndexEntry{Fast Statistical Hypothesis Tests using fastats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE,
  fig.width = 12,
  fig.height = 8,
  dpi = 300
)
```

## Introduction

The `fastats` package provides a suite of highly optimized statistical hypothesis tests for R. By moving the core computational loops to C++ via `Rcpp` and employing aggressive vectorization and zero-allocation techniques, `fastats` offers significant speedups over R's built-in `t.test()` operations, especially for large datasets.

```{r load-package}
library(fastats)
```

## Welch's Two-Sample t-test

The `welch_t_test` function computes this statistic efficiently:

```{r welch-example}
set.seed(42)
x <- rnorm(10000, mean = 0, sd = 1)
y <- rnorm(10000, mean = 0.5, sd = 1.2)

# Perform a test
res_welch <- welch_t_test(x, y, alternative = "two.sided")
res_welch
```

## Paired Two-Sample t-test

The `paired_t_test` function computes the test statistic in a single pass over the data, skipping missing (`NA`) values dynamically without intermediate memory allocation.

```{r paired-example}
pre_treatment <- runif(5000, 10, 20)
post_treatment <- pre_treatment + rnorm(5000, mean = 2, sd = 0.5)

# Adding NA values to demonstrate robust skipping
pre_treatment[c(10, 50, 100)] <- NA

res_paired <- paired_t_test(pre_treatment, post_treatment)
res_paired
```